---
title: "Introduction to jagstargets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to jagstargets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# With the root.dir option below,
# this vignette runs the R code in a temporary directory
# so new files are written to temporary storage
# and not the user's file space.
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
if (identical(Sys.getenv("NOT_CRAN", unset = "false"), "false")) {
  knitr::opts_chunk$set(eval = FALSE)
}
library(R2jags)
library(dplyr)
library(targets)
library(jagstargets)
```

The `jagstargets` package makes it easy to run a single jags model and keep track of the results. [`R2jags`](https://github.com/suyusung/R2jags) fits the models, and [`targets`](https://docs.ropensci.org/targets/) manages the workflow and helps avoid unnecessary computation.

Consider the simple regression model below with response variable `y` and covariate `x`.


$$
\begin{aligned}
y_i &\stackrel{\text{iid}}{\sim} \text{Normal}(x_i \beta, 1) \\
\beta &\sim \text{Normal}(0, 1)
\end{aligned}
$$

We write this model in the JAGS model file below.

```{r}
lines <- "model {
  for (i in 1:n) {
    y[i] ~ dnorm(x[i] * beta, 1)
  }
  beta ~ dnorm(0, 1)
}"
writeLines(lines, "x.jags")
```

A typical workflow proceeds as follows:

1. Prepare a list of input data to JAGS, including vector elements `x` and `y`.
1. Fit the JAGS model using the list of input data.
1. Use the fitted model object to compute posterior summaries and convergence diagnostics.
1. Use the fitted model object to extract posterior draws of parameters and store them in a tidy data frame.
1. If there are other models to compare, use the fitted model object to compute the deviance information criterion (DIC).

`jagstargets` encapsulates this workflow with the [`tar_jags()`](https://docs.ropensci.org/jagstargets/reference/tar_jags.html) function. To use it in a [`targets`](https://docs.ropensci.org/targets/) pipeline, invoke it from the `_targets.R` script of the project.

```{r, echo = FALSE}
# Writes the _targets.R file shown in the next code chunk.
library(targets)
tar_script({
  library(jagstargets)
  options(crayon.enabled = FALSE)
  list(
    tar_jags(
      example,
      jags_files = "x.jags",
      parameters.to.save = "beta",
      data = tar_jags_example_data(),
      stdout = R.utils::nullfile(),
      stderr = R.utils::nullfile()
    )
  )
})
```

```{r, eval = FALSE}
# _targets.R
library(targets)
library(jagstargets)

generate_data <- function(n = 10) {
  true_beta <- stats::rnorm(n = 1, mean = 0, sd = 1)
  x <- seq(from = -1, to = 1, length.out = n)
  y <- stats::rnorm(n, x * true_beta, 1)
  out <- list(n = n, x = x, y = y)
}

# The _targets.R file ends with a list of target objects
# produced by jagstargets::tar_jags(), targets::tar_target(), or similar.
list(
  tar_jags(
    example,
    jags_files = "x.jags",
    parameters.to.save = "beta",
    data = generate_data()
  )
)
```

[`tar_jags()`](https://docs.ropensci.org/jagstargets/reference/tar_jags.html) only *defines* the pipeline. It does not actually run JAGS, it declares the targets that will eventually run JAGS. The specific targets are as follows. Run `tar_manifest()` to show specific details about the targets declared.

```{r}
tar_manifest()
```

Each target is responsible for a piece of the workflow.

* `example_file_x`: Reproducibly track changes to the jags model file.
* `example_data`: Run the code you supplied to the `data` argument of `tar_jags()` and return a dataset compatible with JAGS.
* `example_mcmc_x`: Run the MCMC and return an object of class `rjags` from [`R2jags`](https://github.com/suyusung/R2jags).
* `example_draws_x`: Return a friendly `tibble` of the posterior draws from `example`.
* `example_summaries_x`: Return a friendly `tibble` of the posterior summaries from `example`. Uses [`posterior::summarize_draws()`](https://mc-stan.org/posterior/reference/draws_summary.html)
* `example_dic_x`: Return a friendly `tibble` with each model's DIC and penalty.

The suffix `_x` comes from the base name of the model file, in this case `x.jags`. If you supply multiple model files to the `jags_files` argument, all the models share the same dataset, and the suffixes distinguish among the various targets.

The targets depend on one another: for example, `example_mcmc_x` takes `example_data` as input. [`targets`](https://docs.ropensci.org/targets/) can visualize the dependency relationships in a dependency graph, which is helpful for understanding the pipeline and troubleshooting issues.

```{r, output = FALSE, message = FALSE}
tar_visnetwork(targets_only = TRUE)
```

Run the computation with `tar_make()`.

```{r, output = FALSE}
tar_make()
```

The output lives in a special folder called `_targets/` and you can retrieve it with functions `tar_load()` and `tar_read()` (from [`targets`](https://docs.ropensci.org/targets/)).

```{r}
tar_read(example_summary_x)
```

At this point, all our results are up to date because their dependencies did not change.

```{r}
tar_make()
```

But if we change the underlying code or data, some of the targets will no longer be valid, and they will rerun during the next `tar_make()`. Below, we change the jags model file, so the MCMC reruns while the data is skipped. This behavior saves time and enhances reproducibility.

```{r}
write(" ", file = "x.jags", append = TRUE)
```

```{r}
tar_outdated()
```

```{r}
tar_visnetwork(targets_only = TRUE)
```

```{r, output = FALSE}
tar_make()
```

At this point, we can add more targets and custom functions for additional post-processing. See below for a custom summary target (which is equivalent to customizing the `summaries` argument of `tar_jags()`.)

```{r, echo = FALSE}
# Writes the _targets.R file shown in the next code chunk.
tar_script({
  library(jagstargets)
  options(crayon.enabled = FALSE)
  tar_option_set(memory = "transient", garbage_collection = TRUE)
  list(
    tar_jags(
      example,
      jags_files = "x.jags",
      parameters.to.save = "beta",
      data = tar_jags_example_data(),
      stdout = R.utils::nullfile(),
      stderr = R.utils::nullfile(),
    ),
    tar_target(
      custom_summary,
      posterior::summarize_draws(
        dplyr::select(example_draws_x, -.draw),
        ~posterior::quantile2(.x, probs = c(0.25, 0.75))
      )
    )
  )
})
```

```{r, eval = FALSE}
# _targets.R
library(targets)
library(jagstargets)

generate_data <- function(n = 10) {
  true_beta <- stats::rnorm(n = 1, mean = 0, sd = 1)
  x <- seq(from = -1, to = 1, length.out = n)
  y <- stats::rnorm(n, x * true_beta, 1)
  out <- list(n = n, x = x, y = y)
}

list(
  tar_jags(
    example,
    jags_files = "x.jags",
    parameters.to.save = "beta",
    data = generate_data()
  ),
  tar_target(
    custom_summary,
    posterior::summarize_draws(
      dplyr::select(example_draws_x, -.draw),
      ~posterior::quantile2(.x, probs = c(0.25, 0.75))
    )
  )
)
```

In the graph, our new `custom_summary` target should be connected to the upstream `example` target, and only `custom_summary` should be out of date.

```{r}
tar_visnetwork(targets_only = TRUE)
```

In the next `tar_make()`, we skip the expensive MCMC and just run the custom summary.

```{r, output = FALSE, warning = FALSE}
tar_make()
```

```{r}
tar_read(custom_summary)
```

## Multiple models

`tar_jags()` and related functions allow you to supply multiple models to `jags_files`. If you do, each model will run on the same dataset. Consider a new model, `y.jags`.

```{r}
lines <- "model {
  for (i in 1:n) {
    y[i] ~ dnorm(x[i] * x[i] * beta, 1) # Regress on x^2 instead of x.
  }
  beta ~ dnorm(0, 1)
}"
writeLines(lines, "y.jags")
```

Below, we add `y.jags` to the `jags_files` argument of `tar_jags()`.

```{r, echo = FALSE}
# Writes the _targets.R file shown in the next code chunk.
tar_script({
  library(targets)
  library(jagstargets)
  list(
    tar_jags(
      example,
      jags_files = c("x.jags", "y.jags"),
      parameters.to.save = "beta",
      data = tar_jags_example_data(),
      stdout = R.utils::nullfile(),
      stderr = R.utils::nullfile(),
    ),
    tar_target(
      custom_summary,
      posterior::summarize_draws(
        dplyr::select(example_draws_x, -.draw),
        ~posterior::quantile2(.x, probs = c(0.25, 0.75))
      )
    )
  )
})
```

```{r, eval = FALSE}
# _targets.R
library(targets)
library(jagstargets)

generate_data <- function(n = 10) {
  true_beta <- stats::rnorm(n = 1, mean = 0, sd = 1)
  x <- seq(from = -1, to = 1, length.out = n)
  y <- stats::rnorm(n, x * true_beta, 1)
  out <- list(n = n, x = x, y = y)
}

list(
  tar_jags(
    example,
    jags_files = c("x.jags", "y.jags"),
    parameters.to.save = "beta",
    data = generate_data()
  ),
  tar_target(
    custom_summary,
    posterior::summarize_draws(
      dplyr::select(example_draws_x, -.draw),
      ~posterior::quantile2(.x, probs = c(0.25, 0.75))
    )
  )
)
```

In the graph below, notice how the `*_x` targets and `*_y` targets are both connected to `example_data` upstream.

```{r}
tar_visnetwork(targets_only = TRUE)
```

## More information

For more on [`targets`](https://docs.ropensci.org/targets/), please visit the reference website <https://docs.ropensci.org/targets/> or the user manual <https://books.ropensci.org/targets/>. The manual walks though advanced features of `targets` such as [high-performance computing](https://books.ropensci.org/targets/hpc.html) and [cloud storage support](https://books.ropensci.org/targets/cloud.html).
